Implicit yaddayadda

### 2011-10-28 papers/PLDI2012/implicitization.tex
vire citations de these en francais

All these methods improve the precision of the analysis by keeping the same abstract domain (say, convex polyhedra) but changing the operations applied and their ordering. An alternative is to change the abstract domain, for instance by moving from intervals to octagons or convex polyhedra \cite{DBLP:journals/lisp/Mine06,Mine_PhD}, or the widening operator: for instance one may replace the classical widening on convex polyhedra \cite{CousotHalbwachs78,Halbwachs_PhD} by an improved version \cite{BagnaraHRZ05SCP}, or use widening ``up to'', that is, use linear inequalities syntactically present in the program source code as possible widening steps~\cite{Polka:FMSD:97}.

@@

Classical iterations on the domain of convex polyhedra \cite{Halbwachs_PhD,CousotHalbwachs78} or octagons \cite{Mine_PhD,DBLP:journals/lisp/Mine06} start with $x = 0 \land x = 0$, then continue with $x = y \land 0 \leq x \leq 1$.
The widening operator extrapolates from these two iterations and yields $x = y \land x \geq 0$.
From there, the ``else'' branch at line 5 may be taken; with further widening, $0 \leq y \leq x$ is obtained as a loop invariant, and thus the postcondition computed at line 9 is $x \geq 0 \land y = 0$.
Yet the strongest invariant is $(0 \leq x \leq 51 \land y = x) \lor (51 \leq x \leq 102 \land x+y=102)$, and its convex hull, a convex polyhedron (Fig.~\ref{fig:gopan_reps_invariant}), is
\begin{equation}
y \leq x \land y \leq 102-x \land y \geq 0.\label{eqn:triangle}
\end{equation}
Intuitively, this disappointing result is obtained because widening extrapolates from the first iterations of the loop, but the loop has two different phases ($x \leq 50$ and $x > 50$) with different behaviors, thus the extrapolation from the first phase is not valid for the second.

@@

% TODO: discuter du big step SSA?
% TODO: cite SSA

@@

In cases where the formula is expressed in too rich a logic for the SMT-solver to deal with, a number of workarounds are possible:
\begin{itemize}[itemsep=-\parsep,topsep=0pt,partopsep=0pt]
\item \emph{Linearization}, as per \citet{DBLP:conf/vmcai/Mine06}, which overapproximates nonlinear semantics by linear semantics.
\item Replacing the results of nonlinear operations by ``unknown''.
\end{itemize}
However, such transformations are not yet implemented in our tool.
% TODO: c'est si dur de remplacer * non lineaire par undef?

>>>>

All these methods improve the precision of the analysis by keeping the same abstract domain (say, convex polyhedra) but changing the operations applied and their ordering. An alternative is to change the abstract domain, for instance by moving from intervals to octagons or convex polyhedra \cite{DBLP:journals/lisp/Mine06}, or the widening operator: for instance one may replace the classical widening on convex polyhedra \cite{CousotHalbwachs78} by an improved version \cite{BagnaraHRZ05SCP}, or use widening ``up to'', that is, use linear inequalities syntactically present in the program source code as possible widening steps~\cite{Polka:FMSD:97}.

@@

Classical iterations on the domain of convex polyhedra \cite{CousotHalbwachs78,BagnaraHRZ05SCP} or octagons \cite{DBLP:journals/lisp/Mine06} start with $x = 0 \land x = 0$, then continue with $x = y \land 0 \leq x \leq 1$.
The widening operator extrapolates from these two iterations and yields $x = y \land x \geq 0$.
From there, the ``else'' branch at line 5 may be taken; with further widening, $0 \leq y \leq x$ is obtained as a loop invariant, and thus the postcondition computed at line 9 is $x \geq 0 \land y = 0$.
Yet the strongest invariant is $(0 \leq x \leq 51 \land y = x) \lor (51 \leq x \leq 102 \land x+y=102)$, and its convex hull, a convex polyhedron (Fig.~\ref{fig:gopan_reps_invariant}), is
\begin{equation}
y \leq x \land y \leq 102-x \land y \geq 0.\label{eqn:triangle}
\end{equation}
Intuitively, this disappointing result is obtained because widening extrapolates from the first iterations of the loop, but the loop has two different phases ($x \leq 50$ and $x > 50$) with different behaviors, thus the extrapolation from the first phase is not valid for the second.

@@

@@

In cases where the formula is expressed in too rich a logic for the SMT-solver to deal with, a number of workarounds are possible:
\begin{itemize}[itemsep=-\parsep,topsep=0pt,partopsep=0pt]
\item \emph{Linearization}, as per \citet{DBLP:conf/vmcai/Mine06}, which overapproximates nonlinear semantics by linear semantics.
\item Replacing the results of nonlinear operations by ``unknown''.
\end{itemize}
However, such transformations are not yet implemented in our tool.
\DM{c'est si dur de remplacer * non lineaire par undef?}
